Clique and Constraint Models for Maximum Common (Connected) Subgraph Problems

### 2016-04-19 paper/paper.tex
More gratuitous mucking around with the figures

\usepackage{microtype}                 % typographic perfection
\usepackage{complexity}                % \P, \NP etc
\usepackage{tikz}                      % For pretty pictures
\usepackage{amsmath}                   % \operatorname
\usepackage{amssymb}                   % \triangledown
\usepackage{hyperref}                  % clicky links
\usepackage{cleveref}                  % no need to type Figure etc
\usepackage{gnuplot-lua-tikz}          % graphs
\usepackage[ruled,vlined]{algorithm2e} % algorithms (after cleverref!)
\usepackage{cite}                      % automatically sort mass cites
\usepackage[bottom]{footmisc}          % fix footnote + bottom figure on same page

@@

    \centering
    \input{gen-graph-unconnected-heatmap}
    \caption{The cumulative number of instances solved in under a certain time: on the top, 33\%
        labelled graphs, and in the middle, unlabelled and undirected graphs. On the bottom row, an
        instance-by-instance comparison of the clique model with the best CP model, with 33\%
        labelled graphs (with MAC) on the left, and unlabelled and undirected graphs (with FC) on
    the right.} \label{figure:unconnected-cumulative}
\end{figure}

@@

@@

    \vspace*{1em}
    \centering
    \input{gen-graph-connected-cp}
    \caption{A pairwise comparison of different CP techniques. ?? This looks horrible! Make it look
        better.} \label{figure:connected-cp}
\end{figure}

\subsection{Experimental Comparison of CP Models}\label{mccs-cp-eval}

\Cref{figure:connected-cp} compares the three approaches for ensuring connectedness in CP: by
specific branching (Branch), by specific filtering (Filter), or by combining specific branching and
filtering (Both). We show only results using the best variant for each class, i.e.\ MAC for labelled
graphs, and FC for unlabelled graphs, as the other results are very similar.  On labelled graphs, we
observe that branching often outperforms filtering, and that combining both is rarely much worse
than just doing one or the other, and is often much better, even if on average it is slightly
slower.  On unlabelled graphs, we observe that the three variants have rather similar performance.

\subsection{Ensuring Connectedness in a clique-based approach}\label{mccs-clique}

@@

Finally, we describe the colouring process---an example is shown in \cref{figure:assocrestricted}.
In conventional clique algorithms, a simple greedy sequential colouring is used (possibly with the
help of previous colourings to reduce the computational cost \cite{DBLP:conf/lion/NikolaevBS15}, and
possibly with shortcuts taken for certain vertices \cite{DBLP:journals/cor/SegundoT14}, and possibly
followed by a repair step to improve the colouring \cite{DBLP:conf/walcom/TomitaSHTW10}, or stronger
bounding rules based upon MaxSAT inference
\cite{DBLP:conf/ictai/LiFX13,DBLP:conf/lion/LiJX15,DBLP:journals/cor/SegundoNB15}).  Such colourings
will not give us the required property that vertices in $\mathit{remaining} \cap \mathit{connected}$
come last (so they are selected first by the reverse branching order). Thus we produce two greedy
sequential colourings, first considering the non-branching vertices in $\mathit{remaining} \setminus
\mathit{connected}$, followed by the branching vertices, and concatenate them
(\lineref{makecolours}). This produces a valid colouring, since we do not merge any colour classes
between the two stages, although it may use more colours than a single colouring would\footnote{What
    if we did not guarantee that vertices in $\mathit{remaining} \cap \mathit{connected}$ came last,
    and just used a conventional colouring with the branching rule?  Suppose we had four vertices in
    $\mathit{remaining}$, and produced a colouring $[[v_1, v_2], [v_3], [v_4]]$, and suppose that
    extending $\mathit{solution}$ with $\{ v_1, v_3, v_4 \}$ gives an optimal solution. If $v_4$ was
    not $\mathit{connected}$ yet, we would not branch on that subtree, and the bound could eliminate
    branching on $v_3$ and $v_1$, so we would miss the solution. Thus we cannot simply add the
branching rule without also adapting the combined bound / ordering heuristic.}.

@@

\subsection{Experimental comparison of CP with the Clique-based Approach}\label{mccs-eval}

@@

    \centering
    \input{gen-graph-connected-heatmap}
    \caption{The cumulative number of connected instances solved in under a certain time: on the
        top, 33\% labelled undirected graphs with up to 100 vertices, and in the middle, unlabelled
        and undirected graphs with up to 35 vertices. On the bottom row, an instance-by-instance
        comparison of the association and ``MAC both'' approaches, with 33\% labelled graphs on the
        left, and unlabelled and undirected graphs on the right.} \label{figure:connected-cumulative}
\end{figure}

\cref{figure:connected-cumulative}

@@

@@

\bibliographystyle{splncs}
\bibliography{paper}

>>>>

\usepackage{microtype}                 % typographic perfection
\usepackage{complexity}                % \P, \NP etc
\usepackage{tikz}                      % For pretty pictures
\usepackage{amsmath}                   % \operatorname
\usepackage{amssymb}                   % \triangledown
\usepackage{hyperref}                  % clicky links
\usepackage{cleveref}                  % no need to type Figure etc
\usepackage{gnuplot-lua-tikz}          % graphs
\usepackage[ruled,vlined]{algorithm2e} % algorithms (after cleverref!)
\usepackage{cite}                      % automatically sort mass cites
\usepackage[bottom]{footmisc}          % fix footnote + bottom figure on same page
\usepackage{placeins}                  % \FloatBarrier

@@

    \centering
    \input{gen-graph-unconnected-heatmap}
    \caption{The cumulative number of MCS instances solved in under a certain time: on the top, 33\%
        labelled graphs, and in the middle, unlabelled and undirected graphs. On the bottom row, an
        instance-by-instance comparison of the clique model with the best CP model, with 33\%
        labelled graphs (with MAC) on the left, and unlabelled and undirected graphs (with FC) on
    the right.} \label{figure:unconnected-cumulative}
\end{figure}

@@

?? Connected, not strongly connected. No directed graphs.

@@

    \vspace*{1em}
    \centering
    \input{gen-graph-connected-cp}
    \caption{On top, the cumulative number of MCCS instances solved in under a certain time using
        different CP techniques. Below, instance-by-instance comparisons.}
        \label{figure:connected-cp}
\end{figure}

\subsection{Experimental Comparison of CP Connectedness Techniques}\label{mccs-cp-eval}

\Cref{figure:connected-cp} compares the three approaches for ensuring connectedness in CP: by
specific branching (Branch), by specific filtering (Filter), or by combining specific branching and
filtering (Both). We show only results using the best variant for each class---that is, MAC for
labelled graphs, and FC for unlabelled graphs (the other results are very similar).  On labelled
graphs, we see many instances which are solved by branching but not filtering, and vice versa.
However, combining both is rarely much worse than just doing one or the other, and is often much
better, even if on average it is slightly slower.  On unlabelled graphs, we observe that the three
variants have rather similar performance.

\subsection{Ensuring Connectedness in a Clique-Based Approach}\label{mccs-clique}

@@

Finally, we describe the colouring process---an example is shown in \cref{figure:assocrestricted}.
In conventional clique algorithms, a simple greedy sequential colouring is used (possibly with the
help of previous colourings to reduce the computational cost \cite{DBLP:conf/lion/NikolaevBS15}, and
possibly with shortcuts taken for certain vertices \cite{DBLP:journals/cor/SegundoT14}, and possibly
followed by a repair step to improve the colouring \cite{DBLP:conf/walcom/TomitaSHTW10}, or stronger
bounding rules based upon MaxSAT inference
\cite{DBLP:conf/ictai/LiFX13,DBLP:conf/lion/LiJX15,DBLP:journals/cor/SegundoNB15}).  Such colourings
will not give us the required property that vertices in $\mathit{remaining} \cap \mathit{connected}$
come last (so they are selected first by the reverse branching order). Thus we produce two greedy
sequential colourings, first considering the non-branching vertices in $\mathit{remaining} \setminus
\mathit{connected}$, followed by the branching vertices, and concatenate them
(\lineref{makecolours}). This produces a valid colouring, since we do not merge any colour classes
between the two stages, although it may use more colours than a single colouring would.

(What if we did not guarantee that vertices in $\mathit{remaining} \cap \mathit{connected}$ came
last, and just used a conventional colouring with the branching rule?  Suppose we had four vertices
in $\mathit{remaining}$, and produced a colouring $[[v_1, v_2], [v_3], [v_4]]$, and suppose that
extending $\mathit{solution}$ with $\{ v_1, v_3, v_4 \}$ gives an optimal solution. If $v_4$ was not
$\mathit{connected}$ yet, we would not branch on that subtree, and the bound could eliminate
branching on $v_3$ and $v_1$, so we would miss the solution. Thus we cannot simply add the branching
rule without also adapting the combined bound / ordering heuristic.)

@@

\subsection{Experimental Comparison of the CP and Clique Approaches}\label{mccs-eval}

@@

    \centering
    \input{gen-graph-connected-heatmap}
    \caption{The cumulative number of connected instances solved in under a certain time: on the
        top, 33\% labelled undirected graphs with up to 100 vertices, and in the middle, unlabelled
        and undirected graphs with up to 35 vertices. On the bottom row, an instance-by-instance
        comparison of the association and CP Both approaches, with 33\% labelled graphs on the
        left, and unlabelled and undirected graphs on the right.} \label{figure:connected-cumulative}
\end{figure}

In \cref{figure:connected-cumulative} we compare the clique-based approach to the connected problem
with both CP Both approaches. The trend is broadly similar to the unconnected problem: for labelled
graphs, the clique-based approach is the clear winner, but for unlabelled graphs the clique approach
lags somewhat.

The heatmaps show a more detailed picture. As before, in the unlabelled case, the association
approach is almost never more than an order of magnitude better, and is often much worse. In the
labelled case, however, there are now many instances where the CP approach does much better than the
association approach, despite the association approach remaining much better overall.

?? Can we figure out why? Is it time-to-find or time-to-prove? Assume that this will take up a few
sentences of space.

@@

Adapting clique algorithms: specialised inference, association-aware bound function, .

@@

\FloatBarrier % looks better this way...
\bibliographystyle{splncs}
\bibliography{paper}
